---
title: "Properties that influence wine score"
author: 'Group 28: Aishwin Tikku, Mengran Li, Steven Kwok, Shaoquan Li, Shuning Li'

output:
  pdf_document:
    latex_engine: pdflatex
    number_sections: no
  html_document:
    df_print: paged
fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA, fig.pos = 'H')
options(digits = 4)
```


```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(tidymodels)
library(GGally)
library(car)
library(sjPlot)
library(skimr)
library(kableExtra)
library(knitr)
library(janitor)
```


``` {r echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
Data <- read.csv("dataset28.csv")
```


# Introduction {#sec:Intro}

Wine is an alcoholic drink, produced by various kinds of fermented fruits like grapes, apple or blueberry. There are four kinds of wines, involving white wine, red wine, rose wine and sparkling wine. The difference of wines depends on various kinds of factors, including type of grapes, soil status, and province state. We analysis a data set from the Wine Enthusiast, a famous American wine provider, in this project. Thousands of wines were rated in this data set, where wine with points lower than 80 were filtered.

The aim of our project is discovering properties leading the occurrence of high rated wine, the wine with points larger than 90. The first session visualize the structure, properties, as well as correlations inside the dataset. We, next in order, analysis factors of wine, leading high ranking, thorough the best generalized linear model. Finally, we conclude the entire analysis, as well as discussing what we can do in the future.

# Data Structure and Visualisation {#sec:DSV}
The whole data set has 7 factors and 2000 observations.
```{r echo=FALSE, eval=TRUE, message=FALSE,warning=FALSE}
Data <- Data[, -1]
Data$country <- as.factor(Data$country)
Data$province <- as.factor(Data$province)
Data$title <- as.factor(Data$title)
Data$variety <- as.factor(Data$variety)
Data$winery <- as.factor(Data$winery)
Data$price <- as.integer(Data$price)

# Group by points, those greater than 90 are pass, others are fail
Data <- Data %>% mutate(score = ifelse(points > 90, "Pass", "Fail"))
Data$score <- as.factor(Data$score)
```


```{r his, fig.width=6, fig.height=4, fig.align='center',echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE, fig.cap="\\label{fig:hist} Histogram of Price and Points"}
#### descriptive statistics ####
Data[, c(2, 3)] %>%
  gather(key = "variable", value = "value") %>%
  ggplot() +
  geom_histogram(aes(x = value), fill = "#80C687", color = "#80C687", alpha = 0.8) +
  facet_wrap(~variable, scales = "free")
```

```{r boxplot, eval=TRUE, echo=FALSE, fig.align="center", fig.cap="\\label{fig:box} Boxplot of points by country", message=FALSE, warning=FALSE, out.width='75%', fig.pos='H'}
ggplot(data = Data) +
  geom_boxplot(aes(points, country))
```

```{r echo=FALSE,eval=TRUE,message=FALSE,warning=FALSE}
 Data %>%
  tabyl(country,score)%>%
  adorn_percentages()%>%
  adorn_pct_formatting()%>%
  adorn_ns()
```


```{r echo=FALSE,eval=TRUE,message=FALSE,warning=FALSE}
# chi square test
chisq.test(Data$country, Data$score)
```



```{r dplot, eval=TRUE, echo=FALSE, fig.align="center", fig.cap="\\label{fig:dplot} Density plot by score", message=FALSE, warning=FALSE, out.width='75%', fig.pos='H'}
ggplot(data = Data) +
  geom_density(aes(price, group = score, fill = score))
```



# Methodology {#sec:FDA}
Due to the excessive classification of nations, provinces,varieties and wineries, we need to reduce the dimensionality of the data set and build models separately for discussion.

First of all, we conduct statistics on the number of samples according to the type of wine, and classify the types with a sample number of less than 10 as 'others'.
```{r glm, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
t <- as.matrix(table(Data$variety))

summary(t)

Data$variety <- as.vector(Data$variety)
Data$variety[which(Data$variety %in% row.names(t)[t<11])] = "other"
Data$variety <- as.factor(Data$variety)

```

We will use the generalised linear model to fit a logistic regression model with score as the response, price and variety as the explanatory variable.

```{r glm, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
fit1 <- glm(score ~ price + variety,
            data = Data,
            family = binomial(link = "logit"))
fit1 %>% summary()
```
Notice that only the 'Riesling' in variety is significant at 5% level.
```{r glm, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
Data <- Data %>% mutate(
  variety = case_when(
    variety == "Riesling" ~ "Riesling ",
    !variety %in% c("Riesling ") ~ "others"
  )
)
fit1 <- glm(score ~ price + variety,
            data = Data,
            family = binomial(link = "logit"))

summary(fit1)
```

Similarly, We can use the same method to eliminate interference from too many categories. In order to find this standard, we fit a logistic regression model and check its summary table.
```{r glm, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}

fit2 <- glm(score ~ price + country,
            data = Data,
            family = binomial(link = "logit"))
fit2 %>% summary()
```

We choose variables with p-values less than 0.1. From the summary table, only 'Israel' in country and price have a significant influence on the score.

So we set all countries except 'Israel' as ‘others’ to exclude their influence.
```{r glm, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
Data <- Data %>% mutate(
  country = case_when(
    country == "Israel" ~ "Israel",
    !country %in% c("Israel") ~ "others"
  )
)

fit2 <- glm(score ~ price + country,
            data = Data,
            family = binomial(link = "logit"))

summary(fit2)
```

Next, We select 'Israel' to check 'province' variable. Filter samples whose country is 'Israel' and fit a logistic regression model with score as response, price, province and variety as explanatory vaiables.
```{r glm, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
Israel <- Data %>% filter(country == "Israel")

glm(score ~ price + province,
    data = Israel,
    family = binomial(link = "logit")) %>%
  summary()
```

From the summary table of the above model, it can be seen that the province has no significant effect on the score.

```{r glm, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
fit3 <- glm(score ~ price + variety + country,
            data = Data,
            family = binomial(link = "logit"))

summary(fit3)
```
After comparing the AIC values of the three models, model 3 was finally used for more in-depth research. 

To avoid the overdispersion, we need to compare the value of deviance divided by residual deviance with 1.
```{r glm, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
deviance(fit3)/df.residual(fit3) 
```
The result is less than 1, there is no overdispersion.


Notice that the coefficients of price and variety are positive, which means that expensive wine are more likely to pass(points is greater than 90), while the coefficient of countryIsrael is negative.

And the all coefficients are significant (p-value of 2e-16, 0.00069 and 0.04711), so we qualify the effect of them.

```{r glm, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
plot_model(fit3, show.values = TRUE,
           title = "Odds (price)", show.p = TRUE)
```

```{r glm, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
plot_model(fit4,
           type = "pred", title = "",
           axis.title = c("price", "Probability of pass"))
```

The above figure shows that every unit increase in price, the expensive wines' odds of passing are 1.06 times those of the cheap wine. Similarly, for Riesling and non-Riesling,the Riesling's odds of passing are 4.02 times non-Riesling wine. And for other places produced wine's odds of getting more than 90 points is 0.14 times that of Israel wine.

# Reseult
After establishing the generalised linear model and comparing them, we found that the price is the most significant factor on wine scores. In addition, after analysis, we found that Riesling wine or Israel wine are easier to reach 90 points than other varieties of wine.

# Coclusion and Future Work {#sec:CFW}

For discovering factors affecting the quality of wine more accurately, we can combine multiple data sets with details of Wineries. Different model, additionally, will be trained and tested for selecting the best model. The targeted model will be neural network, a self-learning model which can handle large amount of new data. Lastly, we will construct a system, which can help business searching and identifying good model. This application would be useful for assigning price and prevent purchasing unqualified wines.

# Reference {#sec:Ref}

 Christina, https://cellar.asia/wine/what-is-wine/`



